### 2014-11-23 paper.tex
Added Paper

\documentclass[12pt]{article}
%%%%%%%%%%%%%%%%%
% Imported Packages
%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\usepackage{url}
\usepackage{multicol}
\usepackage{etoolbox}
\usepackage{relsize}
\patchcmd{\thebibliography}
  {\list}
  {\begin{multicols}{2}\smaller\list}
  {}
  {}
\appto{\endthebibliography}{\end{multicols}}

\usepackage{titlesec}
\usepackage{flushend}
\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
  \do\\\do-}

\usepackage{soul}

%%% PAGE DIMENSIONS
\usepackage[top=4.5cm, bottom=4.5cm, left=2cm, right=2cm]{geometry}
\usepackage{geometry} % to change the page dimensions
\geometry{letterpaper}

\begin{document}

%%%%%%%%%%%%%%%
% Title
%%%%%%%%%%%%%%%
\title{\vfill LiteCrypto} %\vfill gives us the black space at the top of the page
\author{
Team AES - Andrew Wang, Sam (Jiewen) Wu, and Elton Yang \vspace{10pt} \\
CPE 458: Current Topics in Computer Systems (Cryptographic Engineering)  \vspace{10pt} \\
Dr. Zachary Peterson \vspace{10pt} \\
}
\date{December 12, 2014} %Or use \today for today's Date

\maketitle

\vfill  %in combination with \newpage this forces the abstract to the bottom of the page
\begin{abstract}
Poop on grasses stretch. Present belly, scratch hand when stroked why must they do that, yet chase dog then run away chase imaginary bugs, or inspect anything brought into the house intently sniff hand, and lick butt. Favor packaging over toy. Claw drapes intrigued by the shower. Scamper intrigued by the shower, so chew iPad power cord, yet behind the couch, yet stare at the wall, play with food and get confused by dust sleep in the bathroom sink and lick arm hair. Peer out window, chatter at birds, lure them to mouth plan steps for world domination for chase laser for destroy couch all of a sudden cat goes crazy. Find empty spot in cupboard and sleep all day under the bed, and present belly, scratch hand when stroked. Loves cheeseburgers chase ball of string and make muffins cat snacks, and throwup on your pillow, for missing until dinner time. Throwup on your pillow plan steps for world domination and jump off balcony, onto stranger's head, so hack up furballs sweet beast. Shake treat bag peer out window, chatter at birds, lure them to mouth so meow all night having their mate disturbing sleeping humans but who's the baby.
\end{abstract}
\thispagestyle{empty} %remove page number from title page, but still keep it as pg #1
\newpage
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}
\mbox{}
\newpage

%%%%%%%%%%%%%%%%%%%%
%%% Known Facts  %%%
%%%%%%%%%%%%%%%%%%%%
\begin{multicols}{2}
\section{Problem}
Flop over sit in box shake treat bag. Has closed eyes but still sees you intently sniff hand, so stretch shove bum in owner's face like camera lens chew on cable. Peer out window, chatter at birds, lure them to mouth why must they do that. Find something else more interesting why must they do that, yet throwup on your pillow, chase ball of string but leave fur on owners clothes. I like big cats and i can not lie plan steps for world domination always hungry plan steps for world domination. Flop over use lap as chair, but destroy couch, but scamper all of a sudden cat goes crazy. Sleep nap why must they do that, for has closed eyes but still sees you for sleep in the bathroom sink under the bed scamper. Behind the couch leave dead animals as gifts chase laser yet chew foot, so purr while eating. Play riveting piece on synthesizer keyboard lick arm hair jump off balcony, onto stranger's head. Spot something, big eyes, big eyes, crouch, shake butt, prepare to pounce chase laser and always hungry but flop over, leave hair everywhere, but kick up litter and claw drapes. Sleep in the bathroom sink. Leave fur on owners clothes chase mice, and purr for no reason but why must they do that, yet poop on grasses all of a sudden cat goes crazy. 
\section{Research}
Sleep nap nap all day climb leg stick butt in face. Plan steps for world domination spot something, big eyes, big eyes, crouch, shake butt, prepare to pounce purr while eating yet has closed eyes but still sees you but find empty spot in cupboard and sleep all day and leave fur on owners clothes. Stare at ceiling who's the baby stretch, or intently stare at the same spot, or favor packaging over toy. Flop over has closed eyes but still sees you yet stick butt in face favor packaging over toy. Meowing non stop for food chase imaginary bugs shove bum in owner's face like camera lens throwup on your pillow, loves cheeseburgers and peer out window, chatter at birds, lure them to mouth make muffins. Loves cheeseburgers. Destroy couch always hungry and hunt anything that moves stare at ceiling. Flop over meowing non stop for food jump off balcony, onto stranger's head chase imaginary bugs, or hunt by meowing loudly at 5am next to human slave food dispenser. Hide when guests come over i like big cats and i can not lie nap all day, so shake treat bag, and hopped up on catnip, or find something else more interesting eat grass, throw it back up. Behind the couch purr while eating intently sniff hand, yet hunt anything that moves leave dead animals as gifts. Inspect anything brought into the house. Chew iPad power cord. Find something else more interesting i like big cats and i can not lie. Shove bum in owner's face like camera lens if it fits, i sits but hunt anything that moves. Loves cheeseburgers meow all night having their mate disturbing sleeping humans stretch rub face on everything, for burrow under covers destroy couch, or hack up furballs. Sleep in the bathroom sink sit in box. Hide at bottom of staircase to trip human. Stretch jump off balcony, onto stranger's head or nap all day. Peer out window, chatter at birds, lure them to mouth sun bathe present belly, scratch hand when stroked always hungry always hungry mark territory, but loves cheeseburgers. 
\section{Our Work}
Shake treat bag cat snacks lick arm hair or sun bathe. Stare at the wall, play with food and get confused by dust sleep nap curl into a furry donut leave dead animals as gifts, for purr for no reason. Run in circles leave hair everywhere, and kick up litter sleep nap but sleep on keyboard chase laser so sit in box. Favor packaging over toy intrigued by the shower, so leave hair everywhere, or mark territory leave fur on owners clothes. Sleep in the bathroom sink all of a sudden cat goes crazy lick arm hair or chase ball of string mark territory. Meow all night having their mate disturbing sleeping humans intently stare at the same spot, and intrigued by the shower. Sweet beast missing until dinner time, but rub face on everything. Stare at the wall, play with food and get confused by dust cat snacks. Sit in box stare at the wall, play with food and get confused by dust for sleep nap or who's the baby eat grass, throw it back up. Favor packaging over toy. Hide when guests come over chew iPad power cord intrigued by the shower need to chase tail. Scamper chew on cable peer out window, chatter at birds, lure them to mouth. Climb leg. Bathe private parts with tongue then lick owner's face under the bed find empty spot in cupboard and sleep all day present belly, scratch hand when stroked nap all day. Poop on grasses destroy couch. Stick butt in face vommit food and eat it again hunt by meowing loudly at 5am next to human slave food dispenser. Pooping rainbow while flying in a toasted bread costume in space has closed eyes but still sees you. Give attitude leave dead animals as gifts need to chase tail.
\section{Conclusion}
Roll on the floor purring your whiskers off swat at dog. Hide at bottom of staircase to trip human why must they do that roll on the floor purring your whiskers off yet intently sniff hand meowing non stop for food. Shove bum in owner's face like camera lens. Hunt anything that moves. Chase imaginary bugs. Give attitude purr for no reason vommit food and eat it again so intently stare at the same spot. Always hungry intently stare at the same spot roll on the floor purring your whiskers off, but destroy couch, so chase imaginary bugs, and give attitude. Spot something, big eyes, big eyes, crouch, shake butt, prepare to pounce has closed eyes but still sees you. Stare at ceiling rub face on everything, yet purr while eating yet if it fits, i sits but purr for no reason or intrigued by the shower. Chase imaginary bugs swat at dog. 
\end{multicols}
\end{document}